---
title: "Visualisation"
output: html_document
date: "2024-07-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(edgeR)
library(limma)
library(ComplexHeatmap)
library(bartools)
library(readxl)
```

## Importing RNA barcode count data
```{r, echo=TRUE, warning=FALSE, message=FALSE}
samplesheet <-readxl::read_excel("/Volumes/bioinf/home/ndalvi/SPLINTR/Data/samplesheet.xlsx")
samplesheet <- as.data.frame(samplesheet)
rownames(samplesheet) <- samplesheet$sample

#barcode DGE
rawdata <- read.delim("/Volumes/bioinf/home/ndalvi/SPLINTR/Results/reverse_complement_2/counts/all_counts_combined.tsv", check.names = FALSE, stringsAsFactors = FALSE)
barcode_data <- DGEList(counts=rawdata[,2:5])
rownames(barcode_data$counts)<- rawdata[,1]

#adding samplesheet information to DGElist object
barcode_data$samples$Sample <- samplesheet$Sample
barcode_data$samples$ORG <- samplesheet$ORG
barcode_data$samples$Batch <- samplesheet$Batch
```
## Data QC and normalisation
### Barcode thresholds
Visualizing the effect of different threshold parameters on barcode data for subsequent normalisation.
```{r, message=FALSE, warning=FALSE}

#removing barcodes with no data 
barcode_data <- barcode_data[rowSums(barcode_data$counts) != 0, ]

print("DGEList dimensions pre- and post-threshold")
thresholdCounts(
  barcode_data,
  type = "absolute",
  threshold = 1,
  minSamples = 1,
  plot = T,
  group = "ORG"
)

print("DGEList dimensions pre- and post-threshold")
thresholdCounts(
  barcode_data,
  type = "absolute",
  threshold = 5,
  minSamples = 1,
  plot = T,
  group = "ORG"
)

print("DGEList dimensions pre- and post-threshold")
thresholdCounts(
  barcode_data,
  type = "absolute",
  threshold = 10,
  minSamples = 1,
  plot = T,
  group = "ORG"
)

print("DGEList dimensions pre- and post-threshold")
thresholdCounts(
  barcode_data,
  type = "absolute",
  threshold = 10,
  minSamples = 2,
  plot = T,
  group = "ORG"
)

print("DGEList dimensions pre- and post-threshold")
thresholdCounts(
  barcode_data,
  type = "relative",
  threshold = 1e-10,
  minSamples = 1,
  plot = T,
  group = "ORG"
)

```

Comments: Continuing QC and normalisation with absolute threshold of at least 5 barcodes in a minimum of 1 sample. 


### Normalisation {.tabset}
```{r message=FALSE, warning=FALSE}
dge.filtered <-
  thresholdCounts(
    barcode_data,
    type = "absolute",
    threshold = 5,
    minSamples = 1,
    plot = F
  )

#normalise samples from sequencing depth to CPM
dge.cpmnorm <- normaliseCounts(dge.filtered, method = "CPM")
```

#### Raw 
```{r}
#plot raw and normalised sequencing depth to get an idea of depth discrepancies between PCR replicates
# raw counts per sample
plotReadCounts(dge.filtered, group = "ORG")
```

#### Normalised
```{r}
# normalised counts per sample
plotReadCounts(dge.cpmnorm, group = "ORG")
```

### Counts per barcode {.tabset}
#### by frequency 
```{r}
# plot detected barcodes ordered by frequency in reference library
plotBarcodeCounts(dge.cpmnorm, log10 = F)

# plot log10 barcode counts
plotBarcodeCounts(dge.cpmnorm, log10 = T)
```

#### by count
```{r}
# order barcodes by count across samples
plotBarcodeCounts(dge.cpmnorm, log10 = F, order = T)

# order barcodes by count across samples with log norm
plotBarcodeCounts(dge.cpmnorm, log10 = T, order = T)
```


## Visualisation 
```{r}
#collapsing replicates
dge.filtered.collapsed <- collapseReplicates(
  dge.filtered,
  group = "ORG",
  method = "mean"
)
```

### Bubble Plot {.tabset}
Note: In "collapsed" plots, SPLINTR replicates have been collapsed into a single variable so that results for a given ORG (eg. ORG38, ORG49) sample represent the mean of all of it's replicates.

#### Individual
```{r}
plotBarcodeBubble(dge.filtered, 
                  proportionCutoff = 5, 
                  labelBarcodes = T,
                  group="ORG")
```

#### Collapsed
```{r}
plotBarcodeBubble(dge.filtered.collapsed, 
                  proportionCutoff = 5, 
                  labelBarcodes = T,
                  group="ORG")
```

### Barcode Plot {.tabset}
#### Individual
```{r}
plotBarcodeHistogram(dge.filtered, topN = 10, alphaLowFreq = 0)
```

#### Collapsed
```{r}
plotBarcodeHistogram(dge.filtered.collapsed, topN = 10, alphaLowFreq = 0)
```

### 
Comments: Overall, barcode composition comparison of both individual replicates (38splA vs 49splA, and 38splB vs 49splB) and collapsed samples (ORG38 vs ORG49) do not show much variation which is expected at this early stage. 


### Principal Components Analysis {.tabset}
#### Sample
```{r}
plotBarcodePCA(dge.filtered, 
               groups = "Sample", 
               ntop = 1000,
               pcs=c(1, 2))

```

#### Organoid
```{r}
plotBarcodePCA(dge.filtered, 
               groups = "ORG", 
               ntop = 1000,
               pcs=c(1, 2))

```

###
Comments: PCA plots show no clustering by sample (PC1) by do show clustering by organoid type (PC2) which accounts for 2% of variance. Further attributes to be considered include: batch, when additional SPLINTR samples are run/sequenced. 

